### 2020-12-05 paper/localpsf.tex
dirac comb dual notation may be confusing, but is needed for correctness

Our approximation is defined by:
\begin{itemize}
	\item how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response})
	\item how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and 
	\item what weighting functions $w_i$ that we use (Section \ref{sec:weighting_functions}). 
\end{itemize}
To briefly summarize, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
\begin{equation*}
	\eta = \mathcal{A} \left(\sum_{i=1}^m \delta_{x_i}\right).
\end{equation*}
The batch of points $x_i$ are chosen so that the support of $\mathcal{A}\delta_{x_i}$ and the support of $\mathcal{A}\delta_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$. Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one vector, we recover many $\varphi_i$. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached. The supports of the impulse responses $\mathcal{A}\delta_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure (described in Section \ref{eq:mean_and_covariance_estimation}) that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions. This procedure is the only part of the paper that requires non-negativity of the integral kernel. The functions $w_i$ are the smoothest functions, in a least-squares sense, that interpolate the points $x_i$. Computing the functions $w_i$ require solving two Poisson PDEs per weighting function, and these PDE solves are done cheaply using multigrid.

@@

Once the $\varphi_i$ and $w_i$ are computed, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format. Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
\begin{equation}
\label{eq:kernel_entries}
	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method [CITE]. Once in H-matrix format, fast H-matrix arithmetic can be used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.

@@

\subsection{Operator access levels}

%In order to solve inverse or optimization problems governed by PDEs, we need to solve linear systems with the Hessian as the coefficient matrix [cite cite]. In order to do this efficiently, we have access to only Hessian-vector products. That is, we have the ability to compute the functions $x \mapsto \mathcal{H} x$ and $x \mapsto \mathcal{H}^T x$ for arbitrary vectors $x$. Therefore

A linear operator may be classified by the way in which we may access the operator.

\begin{itemize}
	\item \emph{Level 1:} A matrix representation of the operator, $\mathbf{A}$, is stored in memory in dense or sparse format.
	\item \emph{Level 2:} We can efficiently compute arbitrary matrix entries $\mathbf{A}_{ij}$ at will, but the matrix $\mathbf{A}$ is not stored in memory.
	\item \emph{Level 3:} Access to $\mathbf{A}$ is only available through an algorithm that computes matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{u} \mapsto \mathbf{A}^T \mathbf{u}$.
\end{itemize}
In large-scale scientific computing,

As discussed in Section \ref{sec:PDE_hessian_motivation}, in inverse problems governed by PDEs, matrix entries of the Hessian are not easily computable since getting one entry would require computing an entire column of the Hessian, then throwing away all but one entry in the column. Computing a single column costs at least two PDE solves. Many such entries are needed to approximate the Hessian in inverse problems with highly informative data (i.e., for Hessians with large numerical rank) [cite cite].

 of the implicit definition of $u(m)$.

Hierarchical matrices (H, H2, HSS, HBS, HODLR), quantized tensor trains, hierarchical interpolative factorization, STRUMPACK (LBNL),

Level 1: not relevant in large-scale problems
State clearly: we do not use Level 1 at all.
This paper: Level 3 -> Level 2
H-matrix: solve on level 2

%\subsection{Existing work}
%
%
%\begin{itemize}
%	\item CITE ESCANDE, others, for a reviews of the subject.
%	\item Resolution analysis seismic papers: try to figure out size of $\varphi_x$ by matvec with random function, post-processing to get autocorrelation length. They did the constant matvec volume thing. Their goal is to get a sense of local length scales that can be resolved in the inverse problem. They do not build preconditioner
%	\item Escande Weiss scattered response paper. They interpolate scattered impulse responses using radial basis functions and study how the geometry of the points affects the approximation accuracy. Matrix-free / getting many points at once is not a concern of them. The applications they consider one has direct access to the impulse response, e.g., through formula for the kernel. 
%	\item Imaging problems: ``matvec'' requires performing an experiment
%	\item pseudodifferential operator equivalence
%	\item Laundry list of applications
%	\item scattered impulse response of escande weiss paper: radial basis functions for approximation in $y$. How to determine if the impulse responses are far enough away from each other?
%	\item my convolution paper: compute impulse responses one at a time, one impulse response per matvec: too slow! Here we compute many impulse responses using only 1 matvec.
%	\item uniformity of smoothness condition
%\end{itemize}

>>>>

Our approximation is defined by:
\begin{itemize}
	\item how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response})
	\item how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and 
	\item what weighting functions $w_i$ that we use (Section \ref{sec:weighting_functions}). 
\end{itemize}
To briefly summarize, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
\begin{equation*}
	\eta = \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
\end{equation*}
Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb. In computations, this amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix. The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one vector, we recover many $\varphi_i$. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached. The supports of the impulse responses $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure (described in Section \ref{eq:mean_and_covariance_estimation}) that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions. This procedure is the only part of the paper that requires non-negativity of the integral kernel. The functions $w_i$ are the smoothest functions, in a least-squares sense, that interpolate the points $x_i$. Computing the functions $w_i$ require solving two Poisson PDEs per weighting function, and these PDE solves are done cheaply using multigrid.

@@

Once the $\varphi_i$ and $w_i$ are computed, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format. Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
\begin{equation}
\label{eq:kernel_entries}
	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$. This allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks [CITE]. Once in H-matrix format, fast H-matrix arithmetic can be used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.

@@

%\subsection{Operator access levels}
%
%
%%In order to solve inverse or optimization problems governed by PDEs, we need to solve linear systems with the Hessian as the coefficient matrix [cite cite]. In order to do this efficiently, we have access to only Hessian-vector products. That is, we have the ability to compute the functions $x \mapsto \mathcal{H} x$ and $x \mapsto \mathcal{H}^T x$ for arbitrary vectors $x$. Therefore 
%
%A linear operator may be classified by the way in which we may access the operator. 
%
%
%\begin{itemize}
%	\item \emph{Level 1:} A matrix representation of the operator, $\mathbf{A}$, is stored in memory in dense or sparse format.
%	\item \emph{Level 2:} We can efficiently compute arbitrary matrix entries $\mathbf{A}_{ij}$ at will, but the matrix $\mathbf{A}$ is not stored in memory.
%	\item \emph{Level 3:} Access to $\mathbf{A}$ is only available through an algorithm that computes matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{u} \mapsto \mathbf{A}^T \mathbf{u}$.
%\end{itemize}
%In large-scale scientific computing, 
%
%
%As discussed in Section \ref{sec:PDE_hessian_motivation}, in inverse problems governed by PDEs, matrix entries of the Hessian are not easily computable since getting one entry would require computing an entire column of the Hessian, then throwing away all but one entry in the column. Computing a single column costs at least two PDE solves. Many such entries are needed to approximate the Hessian in inverse problems with highly informative data (i.e., for Hessians with large numerical rank) [cite cite].
%
% of the implicit definition of $u(m)$. 
%
%Hierarchical matrices (H, H2, HSS, HBS, HODLR), quantized tensor trains, hierarchical interpolative factorization, STRUMPACK (LBNL), 
%
%Level 1: not relevant in large-scale problems
%State clearly: we do not use Level 1 at all.
%This paper: Level 3 -> Level 2
%H-matrix: solve on level 2
%
%%\subsection{Existing work}
%%
%%
%%\begin{itemize}
%%	\item CITE ESCANDE, others, for a reviews of the subject.
%%	\item Resolution analysis seismic papers: try to figure out size of $\varphi_x$ by matvec with random function, post-processing to get autocorrelation length. They did the constant matvec volume thing. Their goal is to get a sense of local length scales that can be resolved in the inverse problem. They do not build preconditioner
%%	\item Escande Weiss scattered response paper. They interpolate scattered impulse responses using radial basis functions and study how the geometry of the points affects the approximation accuracy. Matrix-free / getting many points at once is not a concern of them. The applications they consider one has direct access to the impulse response, e.g., through formula for the kernel. 
%%	\item Imaging problems: ``matvec'' requires performing an experiment
%%	\item pseudodifferential operator equivalence
%%	\item Laundry list of applications
%%	\item scattered impulse response of escande weiss paper: radial basis functions for approximation in $y$. How to determine if the impulse responses are far enough away from each other?
%%	\item my convolution paper: compute impulse responses one at a time, one impulse response per matvec: too slow! Here we compute many impulse responses using only 1 matvec.
%%	\item uniformity of smoothness condition
%%\end{itemize}
